Numerical Investigation of Viscous Flow Instabilities in Multiphase Heterogeneous Porous Media

### 2018-07-11 Papers/ViscousInstabilities/Manuscript_ViscousInstabilities_KC.tex
updates

The geometrical flexibility associated with high-order numerical accuracy of finite element methods (FEM) has proven to be more efficient than FDM to solve fluid flow dynamics in complex geometries. Among FEM-based formulations for porous media flows, the control volume finite element methods~\citep[CVFEM,][]{fung_1992} was designed to guarantee local mass conservation and high-order numerical accuracy as well as being able to use tetrahedral geometry-conforming elements. In traditional CVFEM formulations, pressure and velocity are interpolated using piecewise linear FE basis functions, while material properties and flow conditions (\eg phase saturation, density, temperature, species concentration etc) are represented with CV basis functions~\citep{voller_2009}.

Since geometries are captured by finite elements, constructed control volumes typically extend on each side of the interface which may have different properties. Therefore, average values of coupled velocity-pressure variables are applied across the CVs at the interfaces. This often leads to excessive numerical dispersion especially in highly heterogeneous media. In order to reduce artificial numerical diffusion associated to the upwind weighting of the advected quantity in regions of strong heterogeneity, a discontinuous hybrid finite element finite volume method (DFEFVM) formulation was introduced by \citet{nick_2011b, nick_2011a}. This novel discretisation scheme was designed to simulate flows through discrete fractured rocks in which CVs are divided along the interfaces of different materials. A control-volume distributed (CVD) scheme coupled with mixed FEM (MFEM) was introduced by \citet{edwards_2006}, where flow and rock variables are discretised in CV space whereas pressure and velocity are solved in FEM space. CVD flux-limiting scheme was coupled with high-order convection scheme to reduce anisotropic numerical diffusion in flows through semi-impervious barrier (\ie a region of very low permeability).

>>>>

The geometrical flexibility associated with high-order numerical accuracy of finite element methods (FEM) has proven to be more efficient than FDM to solve fluid flow dynamics in complex geometries. Among FEM-based formulations for porous media flows, the control volume finite element methods~\citep[CVFEM,][]{fung_1992} was designed to guarantee local mass conservation and high-order numerical accuracy as well as being able to use tetrahedral geometry-conforming elements. In traditional CVFEM formulations, pressure and velocity are interpolated using piecewise linear FE basis functions, while material properties and flow conditions (\eg phase saturation, density, temperature, species concentration etc) are interpolated with CV basis functions~\citep{voller_2009}.

Since geometries are captured by FE, constructed CVs typically extend on each side of the interface which may have different properties. Therefore, average values of coupled velocity-pressure variables are applied across the CVs at the interfaces. This often leads to excessive numerical dispersion especially in highly heterogeneous media. In order to reduce artificial numerical diffusion associated to the upwind weighting of the advected quantity in regions of strong heterogeneity, a discontinuous hybrid finite element finite volume method (DFEFVM) formulation was introduced by \citet{nick_2011b, nick_2011a}. This novel discretisation scheme was designed to simulate flows through discrete fractured rocks in which CVs are divided along the interfaces of different materials. A control-volume distributed (CVD) scheme coupled with mixed FEM (MFEM) was introduced by \citet{edwards_2006}, where flow and rock variables are discretised in CV space whereas pressure and velocity are solved in FEM space. CVD flux-limiting scheme was coupled with high-order convection scheme to reduce anisotropic numerical diffusion in flows through semi-impervious barrier  \citep[\ie a region of very low permeability, see also][that solved this particular numerical diffusion problem with the same model used here, originally proposed by \citet{edwards_2006}]{salinas_2018}.
